{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-01 thesis/thesis.tex
add lotsa comment proto-prose thoughts for future work speculation and conclusion

@@

asdfasdasdf
%\appendix
%\include{appendix}

>>>>

% test case design, writing code that actually checks the properties you're interested in
% for example the mutex test, with the if (!xtest) yield check
% then even after you got the right properties, optimizing the state space
% with without_function, ignore_sym, etc, while preserving the guarantee

% envision a framework that uses some code coverage tester to guide choice of pps and/or the user's without_function calls 
% to get that coverage -- although relating flow control coverage directly to concurrency coverage is not straightforward
% framework can also use 'git diff's, like comparing the most recent commit against an already-trusted codebase,
% to figure out how to isolate just the stuff the change affects (memory intersection & data flow resulting from those conflix)
% and then automatically assign some without_functions/etc to test just that commit

% failure injection, testing malloc/etc failures
% eg student condvars which loop around malloc failures can deadlock
% can use retry set reduction in principle to prune the same kind of equivalences

% also here we've got discussion of the exit-UAF buggers where landslide oops verifs the test anyway
% because they're using a free pool and not free()
% need a 3rd thread at least, to expose this (by actually getting corruption)

% things not to forget about
% in the maximal state space, the "tag all evil ancestors" thing justified in my MS thesis figure 31337
% is not necessary, bc if it's the maximal state space, you know that data race PPs would separate the accesses in C1/C2/C3.
% but, actually figuring out if you're in the maximal state space on-the-fly to justify not tagging?
% (but but, does this even matter, is this an optimization at all?)
% i'm actually not even sure this would be sound even on maximal, because of RMW instructions, their 2 evence are inseparable

In the author's experience,
the WBE algorithm (used in Landslide for estimating number of interleavings) consistently underestimates,
being often seen to count (nearly) monotonically up toward the true value as exploration progresses,
while the RE algorithm (used in Landslide for estimating total execution time) can be unstable and erratic,
bouncing wildly among orders of magnitude even in the space of ten or fewer adjacent interleavings.
These flaws are especially visible in the {\tt avl\_fixed} transactional verification results (Table~\ref{tab:tm-verifs}).
% TODO: simpler approach would be to account for like, "T1->T1" prefix subtree likely to be smaller than "T1->T2" prefix subtree
% because T1->T1->... means T1 has 2 fewer events left to schedule and (M+N-2 choose M-2) tends to be smaller than (M+N-1 choose M-1)
% for roughly similar M/N;
% estimator could account for this somehow... not sure how but
% TODO: cite smth about machine learning
% TODO: publish the htm evaluation logs
% state space estimation could probably be drastically improved with machine learning
% accordingly, all the state space estimate logs (ie, quicksand logs) from the tsx chapter are made open-source on so-and-so

% arugment why ML is NOT appropriate for smc in general -- if you try to train it to choose thread interleavings / search ordering,
% even if you're already in not-enough-cpu-time heuristics land, you're biasing it to find only the types of bugs you've already seen
% and destroying its ability to find novelties
% if your ML provides a partial verification, it is impossible for a human to understand what that verification means

@@

Concurrency testing is, in a way, a microcosm for research in general:
an exponentially large problem
% (literally and figuratively for each side of the analogy, respectively)
that invariably thwarts they whoever try to solve it ``perfectly'';
it bears fruit only to one who can find their own human-oriented meaning in partial results and compromises.

asdfasdasdf

% TODO: Get a little bit preachy about how computer programs should be created to help humans,
% not to look as much like scifi as possible, by knowing their limits, asking for help when they get confused, etc.

% ways humans cooperate to make landslide better:
% - test case design incl. preemption sealing
% - abstraction reduction
% - manual inspection of data races

%\appendix
%\include{appendix}
